---
title: A template for the *arxiv* style
authors:
  - name: Juan Ignacio Aberastain
    thanks: Use footnote for providing further information about author (webpage, alternative address)---*not* for acknowledging funding agencies. Optional.
    department: Ingeniería Industrial
    affiliation: Universidad Nacional de Cuyo
    location: Mendoza
    email: mail1
    
  - name: Lucía Cabrera Isquierdo
    thanks: Use footnote for providing further information about author (webpage, alternative address)---*not* for acknowledging funding agencies. Optional.
    department: Ingeniería Industrial
    affiliation: Universidad Nacional de Cuyo
    location: Mendoza
    email: luucabreraisquierdo@gmail.com
    
    - name: Andrea Marina Cortez
    thanks: Use footnote for providing further information about author (webpage, alternative address)---*not* for acknowledging funding agencies. Optional.
    department: Ingeniería Industrial
    affiliation: Universidad Nacional de Cuyo
    location: Mendoza
    email: mail2
    
    - name: Franco Lamotta
    thanks: Use footnote for providing further information about author (webpage, alternative address)---*not* for acknowledging funding agencies. Optional.
    department: Ingeniería Industrial
    affiliation: Universidad Nacional de Cuyo
    location: Mendoza
    email: mail3
    
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
---
## Vectores

Un vector es una estructura de datos que almacena números de doble precisión.

```{r}
mi_vector_a <- c(12,34,12,54,23,12,65,34,12,56,66)
mi_vector_b <- seq(1:16)

mi_vector_a
mi_vector_b

```
## Matrices
Las matrices se parecen a los vectores pero tienen filas ycolumnas. Se alimentan de vectores

```{r}

mi_matriz_c <- matrix(mi_vector_b, nrow=4, byrow=FALSE)

mi_matriz_c

```
Para acceder a un elemento de la matriz uso las filas y columnas entre corchetes
```{r}
mi_matriz_c[2,3]

```
¿Cómo traer la fila 4 completa? 

```{r}
mi_matriz_c[4,]

```
¿Cómo traer una columna?
```{r}
mi_matriz_c[,1]

```
¿Qué hará este comando? Trae toda la matriz menos la fila 2
```{r}
mi_matriz_c [-2,]
```
## Tarea: medir el tiempo que se tarda en generar un vector y una matriz

```{r}
vector_tarea <- seq(1:10000)

start_time <- Sys.time()
vector_tarea
end_time <- Sys.time()


end_time - start_time

```
```{r}
matriz_tarea <- matrix(vector_tarea, nrow=100, byrow=TRUE)

start_time <- Sys.time()
matriz_tarea
end_time <- Sys.time()

end_time - start_time
```
#Tarea de aplicacion 17/4
#Penitencia de Gauss
Mostramos dos formas distintas de sumar las componentes de un vector

```{r}
#Creo el vector
mi_vector<-seq(1:5522)


start_time <- Sys.time()
suma_total <- sum(mi_vector)
end_time <- Sys.time()

end_time - start_time
print(suma_total)
```

```{r} 
start_time <- Sys.time()
suma2_total <- (mi_vector[1]+mi_vector[5522])*(5522/2)
end_time <- Sys.time()

end_time - start_time
print(suma2_total)

```

#Código Fibonacci
Este código suma secuencialmente los números con el anterior. Calculamos adicionalmente el tiempo que tarda en hacerlo
```{r}
start_time<-Sys.time()
a<-0
b<-1
next_num <- a+b
n<-0
while (next_num <= 1000000) {

a<-b
b<-next_num
next_num <- a+b
n<-n+1
end_time<-Sys.time()}
next_num
end_time-start_time
n
```



# Código Burbuja
La siguiente funcion se usa para ordenar numeros de mayor a menor con Bubble sort

```{r}
bubblesort <- function(v) 
{itemCount <- length(v)
    repeat 
    {hasChanged <- FALSE
        itemCount <- itemCount - 1
        for (i in 1:itemCount)
        {if (v[i] > v[i+1]) 
        {t <- v[i]
                v[i] <- v[i+1]
                v[i+1] <- t
                hasChanged <- TRUE}}
        if (!hasChanged) break}
    return(v)}

# Ejemplo de uso
v <- c(123,34,12,54,23,122,65,34,142,56,66)

print(bubblesort(v))

```



# Introduction

Here goes an introduction text

# Headings: first level
\label{sec:headings}

You can use directly LaTeX command or Markdown text. 

LaTeX command can be used to reference other section. See Section \ref{sec:headings}.
However, you can also use **bookdown** extensions mechanism for this.

## Headings: second level

You can use equation in blocks

$$
\xi _{ij}(t)=P(x_{t}=i,x_{t+1}=j|y,v,w;\theta)= {\frac {\alpha _{i}(t)a^{w_t}_{ij}\beta _{j}(t+1)b^{v_{t+1}}_{j}(y_{t+1})}{\sum _{i=1}^{N} \sum _{j=1}^{N} \alpha _{i}(t)a^{w_t}_{ij}\beta _{j}(t+1)b^{v_{t+1}}_{j}(y_{t+1})}}
$$

But also inline i.e $z=x+y$

### Headings: third level

Another paragraph. 


# Examples of citations, figures, tables, references
\label{sec:others}

You can insert references. Here is some text [@kour2014real; @kour2014fast] and see @hadash2018estimate.

The documentation for \verb+natbib+ may be found at

You can use custom blocks with LaTeX support from **rmarkdown** to create environment.

::: {.center latex=true}
  <http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}>
:::

Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  

You can insert LaTeX environment directly too.

\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}

produces

\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}


## Figures

You can insert figure using LaTeX directly. 

See Figure \ref{fig:fig1}. Here is how you add footnotes. [^Sample of the first footnote.]

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

But you can also do that using R.

```{r fig2, fig.cap = "Another sample figure"}
plot(mtcars$mpg)
```

You can use **bookdown** to allow references for Tables and Figures.


## Tables

Below we can see how to use tables. 

See awesome Table~\ref{tab:table} which is written directly in LaTeX in source Rmd file.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

You can also use R code for that.

```{r}
knitr::kable(head(mtcars), caption = "Head of mtcars table")
```


## Lists

- Item 1
- Item 2 
- Item 3
